\LARGE \bf Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation of RGB-D Images

### 2014-02-15 paper.tex
build paper, fixed cite to citep
git-svn-id: https://svn.ais.uni-bonn.de/svn/staff/publications/icra_2014_mueller_nyu@4892 b2511e17-9ac0-488b-bc21-1d6d82e5baa2
Trying to give a better explanation of the learened potentials.
git-svn-id: https://svn.ais.uni-bonn.de/svn/staff/publications/icra_2014_mueller_nyu@4893 b2511e17-9ac0-488b-bc21-1d6d82e5baa2

We evaluate our approach on the public NYU depth V2 segmentation dataset of
indoor scenes.  The dataset comes with a detailed annotation of 1449 
RGB-D images belonging to a wide variety of indoor scenes, categorized into 26
scene classes.  The annotation contains four semantic structural classes:
structure, floor, furniture and prop. There is an additional ``void'' class,
which is used for object boundaries and hard-to-annotate regions. We follow the
literature in excluding these pixels completely from the evaluation.
We optimize our model for \emph{average class accuracy} (the mean of the
diagonal of the confusion matrix), putting more emphasis on the harder
classes of props and furniture, which have smaller area than structure and
floor. The dataset is split into 795 images for training and 654 images for testing.

Our approach is implemented using our \textsc{PyStruct}
library\footnote{\url{http://pystruct.github.io}}.  We use \textsc{scikit-learn}~\citep{pedregosa2011scikit}
for $k$-means clustering and the SVM baseline.

All hyper-parameters were adjusted using 5-fold cross-validation on the training set. The
hyper parameters of the random forests were found using the \texttt{hyperopt}
framework~\cite{bergstra2011algorithms}. For the CRF model, the only
hyper-parameters are related to the superpixel segmentation, and the single
hyper-parameter $C$ of the structural SVM formulation. These were adjusted
using grid search. 
We found 500 superpixels per image to work best, which allow for a maximum
possible performance of 95\% average class accuracy on the validation set.
Training of the random forests took about 15\,minutes on a NVIDIA GeForce GTX
Titan. Training the structural SVM took 45\,minutes on a Xeon X5650 CPU.
Prediction using only the random forest takes 33\,ms on average, while segmentation and prediction
using the structural SVM approximately take an additional 500\,ms.

\subsection{Results}\seclabel{results}
\Tabref{results} compares different components of our approach with the literature.
Please note that we \emph{first} designed our final model, using only the
validation data. We now report accuracies of simpler models for reference,
but these results were not used for model selection. To separate
the influence of loss-based training and the spatial reasoning of the CRF,
we also train a usual support vector machine (SVM) on the unary potentials for comparison.

@@

A visualization of the impact of each processing step can be found in
\Figref{comparison}, which shows prediction results on the test set.  The four
prediction methods correspond to the rows of \Tabref{results}.  The difference
between the SVM and CRF approaches are clearly visible, with the CRF producing
results that are very close to the ground truth in several complex scenes.

We found that our approach improves results most for scenes with a clear geometric structure,
which is not surprising. We see that evidence from the random forest is often very noisy, and biased
away from the ``props'' class. While the unstructured SVM can correct somewhat
for the class imbalance, it has no way to make larger areas consistent, which
the CRF can.
On the other hand, performance of the CRF deteriorates slightly on very crowded
scenes with a mixture of small furniture and prop objects, as
can be seen in the two right-most images. In these scenes, depth information is
often noisy, and it is hard to make geometric statements on the superpixel
level. As the input from the random forest is also often of low quality for
crowded scenes, the CRF has little chance to recover.

@@

\Figref{weights} visualizes two of the learned potential functions.
Higher values correspond to favored configurations. One can see that the
vertical alignment potential expresses that the floor is much more likely to be below
other classes. It also encodes the fact that props
rest on furniture, but not the other way around.
The potential of the depth feature encodes, for example, that the ground
is usually behind the other classes, while furniture is in front of structures,
such as the wall.
Interestingly, the potential functions are not anti-symmetric, and forcing them to be so
degrades the results. This suggests that the direction of connecting edges, going from
the top left to the bottom right in the image, is also exploited by the potentials.

@@

    \includegraphics[width=\linewidth]{images/00845_gt.png}&%
    \includegraphics[width=\linewidth]{images/00781_gt.png}&%
    \includegraphics[width=\linewidth]{images/01331_gt.png}&%
    \includegraphics[width=\linewidth]{images/00118_gt.png}&%
    \includegraphics[width=\linewidth]{images/01203_gt.png}\\
    \multicolumn{5}{@{}c@{}}{Ground Truth}\vspace{3mm}\\
    \multicolumn{5}{@{}c@{}}{%
        \includegraphics[]{legend.pdf}
    }
    \end{tabu}
}\caption{%
Qualitative evaluation of the CRF\@.
The first three images illustrate errors in the original prediction that can be corrected, while
the second two images illustrate failure modes. Pixels marked as void
are excluded from the evaluation. See the \secref{results} for details.
\figlabel{comparison}}
\end{figure*}

\section{SUMMARY AND DISCUSSION}
\enlargethispage{5mm}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model on the output of an
efficient GPU implementation of random forest, and model spatial neighborhood
using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function
that is learned using the SSVM approach. By explicitly modeling 3D relations in
a fully learned framework, we improve the state-of-the-art on the NYU V2
dataset for semantic annotation of structure classes.

While our approach allows modeling of spatial relations, these are limited to
local interactions. In future work, these interactions could be extended to
larger areas using latent variable models or higher order potentials~\citep{ladicky2009associative}.
Another possible line of future work is to combine our approach with a more
task-specific one, directly including support plane assumptions into the model,
as done by \citet{SilbermanECCV12}. Finally, we could also combine our
single-frame approach with the approach of \citet{stueckler2013}, which fuses
individual views in 3D to exploit temporal coherence.

While this work did not explicitly address real time application, the random
forest implementation that we build upon allows for real-time
processing~\citep{stueckler2013}. The SLIC superpixel algorithm can also be
implemented on GPU in real-time, as was demonstrated by \citet{ren2011gslic}, and similarly
the normal features we use also have
real-time capabilities~\citep{holz_2011_robocup}.
Finally, fusion move inference for our model is very efficient for our model,
opening up the possibility to implement our approach entirely in real time.

>>>>

We evaluate our approach on the public NYU depth V2 segmentation dataset of
indoor scenes.  The dataset comes with a detailed annotation of 1449 
RGB-D images belonging to a wide variety of indoor scenes, categorized into 26
scene classes.  The annotation contains four semantic structural classes:
structure, floor, furniture and prop. There is an additional ``void'' class,
which is used for object boundaries and hard-to-annotate regions. We follow the
literature in excluding these pixels completely from the evaluation.
We optimize our model for \emph{average class accuracy} (the mean of the
diagonal of the confusion matrix), putting more emphasis on the harder
classes of props and furniture, which have smaller area than structure and
floor. The dataset is split into 795 images for training and 654 images for testing.
%
Our approach is implemented using our \textsc{PyStruct}
library\footnote{\url{http://pystruct.github.io}}.  We use \textsc{scikit-learn}~\citep{pedregosa2011scikit}
for $k$-means clustering and the SVM baseline.

All hyper-parameters were adjusted using 5-fold cross-validation on the training set. The
hyper parameters of the random forests were found using the \texttt{hyperopt}
framework~\citep{bergstra2011algorithms}. For the CRF model, the only
hyper-parameters are related to the superpixel segmentation, and the single
hyper-parameter $C$ of the structural SVM formulation. These were adjusted
using grid search. 
We found 500 superpixels per image to work best, which allow for a maximum
possible performance of 95\% average class accuracy on the validation set.
Training of the random forests took about 15\,minutes on a NVIDIA GeForce GTX
Titan. Training the structural SVM took 45\,minutes on a Xeon X5650 CPU.
Prediction using only the random forest takes 33\,ms on average, while segmentation and prediction
using the structural SVM approximately take an additional 500\,ms.

\vspace{-2mm}
\subsection{Results}\seclabel{results}
\vspace{-1mm}
\Tabref{results} compares different components of our approach with the literature.
Please note that we \emph{first} designed our final model, using only the
validation data. We now report accuracies of simpler models for reference,
but these results were not used for model selection. To separate
the influence of loss-based training and the spatial reasoning of the CRF,
we also train a usual support vector machine (SVM) on the unary potentials for comparison.

@@

A visualization of the impact of each processing step can be found in
\Figref{comparison}, which shows prediction results on the test set.  The four
prediction methods correspond to the rows of \Tabref{results}.  The difference
between the SVM and CRF approaches are clearly visible, with the CRF producing
results that are very close to the ground truth in several complex scenes.
%
We found that our approach improves results most for scenes with a clear geometric structure,
which is not surprising. We see that evidence from the random forest is often very noisy, and biased
away from the ``props'' class. While the unstructured SVM can correct somewhat
for the class imbalance, it has no way to make larger areas consistent, which
the CRF can.
On the other hand, performance of the CRF deteriorates slightly on very crowded
scenes with a mixture of small furniture and prop objects, as
can be seen in the two right-most images. In these scenes, depth information is
often noisy, and it is hard to make geometric statements on the superpixel
level. As the input from the random forest is also often of low quality for
crowded scenes, the CRF has little chance to recover.

@@

\Figref{weights} visualizes two of the learned potential functions.
Higher values correspond to favored configurations. We did not force potentials
to be symmetric or anti-symmetric, which makes interpretation of the figures a
bit harder, but increases performance.
Edges are constructed to go from top left to bottom right. Potentials below the
diagonal are those for an edge going from a label given by the column to the
one given by the row, while the ones above the diagonal are for the opposite
direction.  For vertical alignment, one would therfore expect to find
anti-symmetric potentials. However the left-right direction
seems to also contain useful information, breaking the symmetry.
One can see that the vertical alignment potential expresses that the floor is
much more likely to be below other classes. It also encodes the fact that props
rest on furniture, but not the other way around.  The potential of the depth
feature encodes, for example, that the ground is usually behind the other
classes, while furniture is in front of structures, such as the wall.

@@

    \includegraphics[width=\linewidth]{images/00845_gt.png}&%
    \includegraphics[width=\linewidth]{images/00781_gt.png}&%
    \includegraphics[width=\linewidth]{images/01331_gt.png}&%
    \includegraphics[width=\linewidth]{images/00118_gt.png}&%
    \includegraphics[width=\linewidth]{images/01203_gt.png}\\
    \multicolumn{5}{@{}c@{}}{Ground Truth}\vspace{3mm}\\
    \multicolumn{5}{@{}c@{}}{%
        \includegraphics[]{legend.pdf}
    }
    \end{tabu}
}\caption{%
Qualitative evaluation of the CRF\@.
The first three images illustrate errors in the original prediction that can be corrected, while
the second two images illustrate failure modes. Pixels marked as void
are excluded from the evaluation. See the \secref{results} for details.
\figlabel{comparison}}
\end{figure*}
\section{SUMMARY AND DISCUSSION}
\enlargethispage{5mm}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model on the output of an
efficient GPU implementation of random forest, and model spatial neighborhood
using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function
that is learned using the SSVM approach. By explicitly modeling 3D relations in
a fully learned framework, we improve the state-of-the-art on the NYU V2
dataset for semantic annotation of structure classes.
%
While our approach allows modeling of spatial relations, these are limited to
local interactions. In future work, these interactions could be extended to
larger areas using latent variable models or higher order potentials~\citep{ladicky2009associative}.
Another possible line of future work is to combine our approach with a more
task-specific one, directly including support plane assumptions into the model,
as done by \citet{SilbermanECCV12}. Finally, we could also combine our
single-frame approach with the approach of \citet{stueckler2013}, which fuses
individual views in 3D to exploit temporal coherence.
%
While this work did not explicitly address real time application, the random
forest implementation that we build upon allows for real-time
processing~\citep{stueckler2013}. The SLIC superpixel algorithm can also be
implemented on GPU in real-time, as was demonstrated by \citet{ren2011gslic}, and similarly
the normal features we use also have
real-time capabilities~\citep{holz_2011_robocup}.
Finally, fusion move inference for our model is very efficient for our model,
opening up the possibility to implement our approach entirely in real time.
